Molecular Gas Dynamics of the strongly-lensed wet-merger RXS J1131$-$1231 at $z$=0.654

### 2016-08-04 WRITEUP/ms.tex
added RP as co author
finised revising abstract & concl.

\documentclass[]{emulateapj}
\usepackage{amsmath}
\input{preamble}
\slugcomment{To be submitted to the ApJ}
 \makeatletter
% \renewcommand\normalsize{\@setfontsize\normalsize\@xpt{11.56}}
\renewcommand\normalsize{\@setfontsize\normalsize{10.56}{11.4}}
\makeatother

\citestyle{aa}
\shorttitle{Molecular Gas in the Lensed Wet-Merger RXJ1131$-$1231 at $z$\,=\,0.65}
\shortauthors{Leung \& Riechers}

@@

\title{Molecular Gas Dynamics of the strongly-lensed wet-merger RXS J1131$-$1231 at $z$=0.654}
\author{T. K. Daisy Leung and Dominik A. Riechers}
\affil{Department of Astronomy, Space Sciences Building, Cornell University,
Ithaca, NY 14853, USA; \myemail}

@@

\begin{abstract}
We present interferometric observations of \bco and \cco emission toward the ``Einstein ring''
RXS J1131$-$1231 at $z$\,=\,0.654, obtained with the \pdbi and \carma. 
Our observations detect a CO-emitting source near RXJ1131, confirming the presence of an optically 
faint companion reported in previous studies. 
Based on our dynamical lens modeling of the spatially and spectrally resolved \bco emission, 
we reconstruct a symmetric velocity gradient in the source plane and 
find an intrinsically symmetric double-horned line profile, which are suggestive of a 
rotating disk morphology, consistent with previous optical studies.
We find a lensing-corrected gas mass of $M_{\rm gas}$\ssim1.4\E{10}\,\Msun in the host galaxy and 
\ssim2\E{9}\,\Msun in the companion (gas mass ratio $\sim$7:1). 
The CO disk is extended over $\sim$6\,kpc, which yields a dynamical mass of $M_{\rm dyn}$\ssim8\E{10}\,\Msun
for the host galaxy. This corresponds to a gas mass fraction of $\sim$19\%, confirming its gas-rich nature.
This fraction also confirms the trend of decreasing gas fraction since $z$\ssim1, which is responsible for the decline of
the cosmic \SF rate and black hole accretion rate densities.
Fitting dust SED models, we derive a lensing-corrected dust mass of $M_{\rm dust}$ = 4\E{8}\,\Msun, a 
lensing-corrected $L_{\rm IR}$\ssim1.5\E{12}\,\Lsun, and a SFR$_{\rm FIR}$\eq120\pmm63 \sfrU, 
confirming the dusty nature of the merger.
A comparison of the gas and stellar distributions reveals active starbursts embedded 
within the molecular gas reservoir which is likely a result from merger-driven gas inflow.
The nature of RXJ1131 is consistent with the picture predicted by model simulations, in which it will 
likely retain its disk component and evolve into small bulge disk galaxy upon merging.
Our study presents the first resolved CO imaging at intermediate redshift, 
which allows us to provide the first dynamical constraint on the Magorrian relation at intermediate redshift.
\end{abstract}

\keywords{ISM: molecular --
          infrared: galaxies --
          galaxies: mergers --
          galaxies: starburst --
          galaxies: evolution}

%--------------------------------------------------------------------------
%                                Introduction
%--------------------------------------------------------------------------
\section{Introduction}
Studies of galaxy evolution have been focusing on the interplay between baryonic physics/processes
(\eg feedback and gas dissipation) and \SF and active galactic nucleus (AGN) activity across cosmic epochs.
It is not well-understood when and how baryons of present-day galaxies were assembled, 
but it is clear that both the co-moving \SF rate and black hole accretion rate densities
increased substantially since $z$\,$>\,3$ and reached their climax at $z$\ssim$2$, followed by 
a rapidly decline toward $z$\ssim$0$ \citep[\eg][]{Hopkins06a, Madau14a}. 
% Studying the cold gas at the redshift range 0.2 $<$\,$z$\,$<$\,1 of paramount importance to ravel the cause for this decline.
A leading explanation for this decline is related to the evolution of molecular gas content and 
\SF efficiency \citep[]{Erb06a, CW13, Walter14a},
but confirmation from direct molecular gas measurements at intermediate redshift 
$(0.2\,<\,z\,<\,1)$ is still limited both in number and in the variety of galaxy populations probed --- 
only unresolved CO observations have been carried out for $\sim$$70$ ULIRGs up to date \citep{Combes11a, Combes13a}.

% In the current paradigm, it is posited that every nearby, massive galaxy host a supermassive black hole (SMBH), which
% grows by accretion of the cold molecular gas from its host galaxy.
Meanwhile, empirical scaling relations such as the Magorrian relation \citep{Magorrian98a}
--- a tight correlation between black hole mass and stellar bulge mass ---
have been established locally, implying a self-regulating co-eval growth between local SMBHs and their host galaxies.
Attempts to extend this relation out to higher redshifts, beyond the peak epoch 
of \SF and AGN activity, have been made in recent years.
It has been reported that the $M_{\rm BH}$$/$$M_{\rm bulge}$ ratio of high-$z$ submillimeter galaxies (SMGs) falls below 
the local relation \citep[\eg][]{Borys05a,Alexander08a} but that of high-$z$ quasar host galaxies are found 
lying above 
the relation \citep[\eg][]{McLure06a,Peng06a}.
Stellar mass of quasar host galaxies at $z$\,$>$\,4 constrained by their CO-based dynamical mass 
are also too small compared to that based from the local relation \citep{Walter04a, Riechers08a,Coppin08a}. 
While it has become clearer that high-$z$ galaxies deviate from the local relation, and among their galaxy populations, 
such a picture is still unclear  at intermediate redshift.
This owes to the difficulties in measuring stellar emission in the host galaxy under the presence of a bright quasar, or alternatively
owing to the lack of spatially resolved CO observations to provide dynamical constraints on the stellar mass. 
With gravitational lensing, emission of the background quasar and its host galaxy 
is magnified and stretched out to different extents, and is being projected onto different locations in the image plane.
This in effect increases the surface brightness contrast between their emission and allows for the
relatively uncontaminated measurements of their emission.

\defcitealias{Sluse03a}{S03}
The ``Einstein ring'' RXS J113151.62$-$123158 (hereafter RXJ1131) 
at $z_\textrm{s, QSO}$\eq$0.685$ \citep[hereafter S03]{Sluse03a} is a particularly well-suited source for
studying the evolution of molecular gas properties in quasar host galaxies and the 
connection between SMBHs and their host galaxies at intermediate redshift
given its unique lensing configuration.
The stellar emission in the host galaxy of RXJ1131 is lensed into
an Einstein ring of radius $1$\farcs$83$
that is clearly separated from the quadruply imaged quasar emission \citep[hereafter C06]{Claeskens06a}.
We have thus carried out a detailed study of RXJ1131 as presented in this paper.
RXJ1131 is strongly lensed by an elliptical galaxy at $z_\textrm{L}$\eq$0.295$ \citepalias{Sluse03a}.
The moderate mass black hole $M_{\rm BH}$\eq$8$\E{7}\Msun \citep{Sluse12a} %based on H beta line
residing in RXJ1131 is rotating with a high spin parameter \citep[$a$\,$\sim$\,$0.9$;][]{Reis14a}, and
is radiating at an intrinsic bolometric luminosity of $L_{\rm bol, X}$\eq$1.3$\E{45}\, ergs\,s\pmOne \citep{Pooley07a}.

This paper is structured as follows.
In \Sec{obs}, we outline the details of the observations and our data reduction process.
In \Sec{HST}, we describe our astrometric corrections applied to align an {\it HST} image with our VLA radio image.
In \Sec{results}, we report measurements of the \bco and \cco lines and a compilation of photometric data from optical to radio wavelengths.
In \Sec{anal}, we present dynamical lens modeling of the \bco data, SED modeling of the thermal dust continuum, 
and the resulting physical properties inferred for RXJ1131.
In \Sec{diss}, we discuss the merger stage of RXJ1131, 
the scenario of a recoiling black hole, and the $M_{\rm BH}$$-$\mdyn relation.
Finally, we summarize the main results and present our conclusions in \Sec{sum}.
We use a concordance $\Lambda$CDM cosmology throughout this paper, with
parameters from the WMAP9 results:
$H_0$ = $69.32$ \kms Mpc\pmOne, $\Omega_{\rm M}$ = $0.29$, and
$\Omega_{\Lambda}$ = $0.71$ \citep{Hinshaw13a}.
% same as LR16a: i.e. use last column (WMAP+eCMB+BAO+H0) of Tab 4 in Hinshaw+13 because that's what astropy use for WMAP9 as cosmo
\defcitealias{Claeskens06a}{C06}

%--------------------------------------------------------------------------
%                          Observations details
%--------------------------------------------------------------------------
\section{Observations} \label{sec:obs}
\subsection{PdBI \bco}
Observations of the \bco rotational line
($\nu_{\rm rest}$\,=\,$230.5379938$ GHz) redshifted to $\nu_{\rm obs}$\,=\,$139.0$\,GHz
were carried out using IRAM \pdbi (PdBI; Program ID: S14BX; PI: D.
Riechers). Two observing runs were carried out on 2014 December 06 and 2015
February 05 under good weather conditions in the C and D array configurations,
respectively. The 2 mm receivers were used to cover the redshifted \bco line
and the underlying continuum emission, employing a correlator setup that provides
an effective bandwidth of 3.6 GHz (dual polarization) and a spectral resolution of 10.0 MHz ($\sim$
$21.5$ \kms). This resulted in 3.75 hours of cumulative six antenna-equivalent on-source
 time after discarding unusable visibility data.
The nearby quasars B1127$-$145 and B1124$-$186 were observed every 22 minutes
for pointing, secondary amplitude, and phase calibration, and B1055$+$018 was
observed as the bandpass calibrator for both tracks.
MWC349 and 3C279 were observed as primary flux calibrators for the C and D
array observations, respectively, yielding $\lesssim$15\% calibration accuracy.

The \ncode{gildas} package was used to reduce and analyze the visibility data.
The calibrated visibility data were imaged and deconvolved using the CLEAN algorithm with ``natural''
weighting. This yields a synthesized clean beam size of 4$\farcs$44 $\times$ 1\farcs95 (PA = 13\degr).
The final rms noise is $\sigma$\,=\,1.45\,mJy\,beam\pmOne
over 10 MHz (21.5\,\kms). The continuum image at $\nu_{\rm cont}\sim$139\,GHz
is created by averaging over 3.16\,GHz of line-free bandwidth. This
yields an rms noise of 0.082 mJy\,beam$^{-1}$. % see README.md in 04Sep15

@@

The \ncode{miriad} package was used to calibrate the visibility data.
The calibrated visibility data were
imaged and deconvolved using the CLEAN algorithm with ``natural'' weighting. This yields a synthesized clean
beam size of 3\farcs2 $\times$ 1\farcs9 (PA\,=\,8\degr) for the lower sideband
image cube. The final rms noise is $\sigma$ = 13.3 mJy beam$^{-1}$
over a channel width of 25\,MHz. An rms noise of
$\sigma$\,=\,0.83\,mJy\,beam\pmOne is reached by averaging over the
line-free channels.

\subsection{VLA (Archival)} %DONE
Our analysis also uses archival data of the 5\,GHz
radio continuum obtained with the
Very Large Array (VLA; Program ID: AW741; PI: Wucknitz).
Observations were carried out on 2008 December 29 under excellent weather
conditions in the A array configuration for a total of $\sim$7 hours on-source time. The C-band receivers were used with a continuum mode setup,
providing a bandwidth of 50 MHz for the two IF bands with full polarization.
The nearby radio quasar J1130$-$149 was observed every 10 minutes for
pointing, amplitude, and phase calibration, J1331$+$305 was observed as the
primary flux calibrator, and J0319$+$415 was observed as the bandpass
calibrator, yielding $\sim$10\% calibration accuracy.
We used \ncode{aips} to calibrate the visibility data.
The calibrated visibility data were imaged and deconvolved using
the CLEAN algorithm using robust\,=\,0. This yields a synthesized clean
beam size of 0$\farcs$49 $\times$ 0\farcs35 (PA\,=\,0.18$\degr$) and a final
rms noise of $\sigma$ = 13 $\mu$Jy beam\pmOne.

\section{HST astrometry} \label{sec:HST}
We obtained an {\it HST} image taken with
the ACS/Wide Field Camera 
using the F555W filter ($V$-band)
from the
Hubble Legacy Archive\footnote{Based on observations
made with the NASA/ESA Hubble Space Telescope, and obtained from the Hubble
Legacy Archive, which is a collaboration between the Space Telescope Science
Institute (STScI/NASA), the Space Telescope European Coordinating Facility
(ST-ECF/ESA) and the Canadian Astronomy Data Centre (CADC/NRC/CSA).}. 
The details of the observations can be found
in \citetalias{Claeskens06a}.
We adopt the VLA 5\,GHz map as the
reference coordinate frame to align the optical ($V$-band) image.
We shift the latter to the east by 0\farcs5963 in R.A. and $+$0\farcs8372 in
Dec., which is consistent with the typical astrometric precision (1$^{\prime\prime}-$2$^{\prime\prime}$) of
images from the Hubble Legacy Archive\footnote{http://hla.stsci.edu/hla\_faq.
html}. This astrometric correction is critical to avoid artificial spatial
offsets between different emitting regions and to carry out our lens modeling,
in which the absolute position of the foreground lensing galaxy is guided by its
coordinates in the high-resolution optical image.
The VLA image is calibrated using a well-monitored phase
calibrator, with absolute positional accuracy of $\sim$2 mas.
For this reason, the absolute alignment between the VLA image and other
interferometric images reported in this paper are expected to have an astrometric
precision better than 0\farcs1, modulo uncertainties related to the SNR and phase
instability.

%--------------------------------------------------------------------------
%                                Results
%--------------------------------------------------------------------------
\section{Results} \label{sec:results}
\subsection{\bco Emission} \label{sec:CO21} %DONE
We detect \bco line emission toward the background source
at $\gtrsim$\,27$\sigma$ significance, confirming the redshift at $z_{\rm CO}$ =
0.65370\,$\pm$\,0.0005. The emission is spatially and dynamically resolved
with a highly asymmetric double-horned line profile
as shown in \Fig{CO21spec}. Fitting a double Gaussian results in peak
flux densities of 75.3\pmm2.6 mJy and 24.0\pmm2.0 mJy, and a FWHM of
179\pmm9 \kms and 255\pmm28 \kms for the two components, respectively. The peaks are separated by
$\Delta v_{\rm sep}$ = 400\pmm12\,\kms. The total integrated line flux is 24.1\,$\pm$\,2.3 Jy \kms. % see 15May16/

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.45\textwidth]{../Figures/SpecCO21_twinx.eps}
\caption{ Spectrum of \bco emission toward RXJ1131. The velocity scale
is with respect to $z$\,=\,0.6537, which is approximately the line center
considering the asymmetry resulting from differential lensing.
A detailed discussion of this effect is presented in
\Sec{differential}.
 \label{fig:CO21spec}}
\end{figure}

% moment 0 map, highest SNR: CASA chan=125~159 <=> GILDAS 126-160 <=> python 125-160
% - sigma = 0.305 Jy km/s/Beam
%   - note that theoretical sigma is lower = 1.5 * sqrt(160-126+1) * 21.5 ~ 0.2 Jy km/s /beam, due to higher noise in some channels with emission
We construct the zeroth moment map, a red/blue channel map, and
the first and second moment maps in \Fig{CO21mom}
using the $uv$-continuum subtracted data cube over a velocity range of
$\Delta v$ $\sim$ 750\,\kms.
The higher-order moment maps are created using
unbinned channel maps with 3$\sigma$ clipping.
The peak flux density is 8.12\pmm0.30 Jy\,\kms\,beam\pmOne
in the intensity map.

The deconvolved source size FWHM from fitting a two-dimensional Gaussian
is 5\farcs1\pmm0\farcs72 $\times$ 3\farcs72\pmm0\farcs66,
and thus, the emission is resolved over $\sim$2.2 beams.
While the lensed emission is not strictly distributed as a two-dimensional
Gaussian;
the fit recovers the line intensity enclosed by the emitting
region, we therefore take this as an estimate on the extent of the lensed
emission. On the other hand, if we assume that the spatial distribution of
the lensed molecular gas emission is similar to that in the optical to \nir
wavelengths, the lensed emission would be more accurately described by an
annulus, enclosing the partially complete ``Einstein ring'' and
the lensed knots (see \Fig{CO21mom}).

\begin{figure*}[!htbp]
\hspace{0.5em}
\includegraphics[trim=0 10 15 0, clip, width=0.452\textwidth]{../Figures/{F555WCO21_mom0_single.invertedgray}.eps}
\includegraphics[trim=5 -10 0 0, clip, width=0.472\textwidth]{../Figures/F555W_REDBLUE.eps} 
\\
\includegraphics[trim=25 16 0 10, clip, width=0.85\textwidth]{../Figures/CO_highOmom_CLIP5sigma}
\vspace{0.1em}
\caption{Top left: overlay of the velocity-integrated \bco emission on the archival {\it HST} $V$-band (F555W)
image. 
Top right: same as top left, except the contours are color-coded to represent the red- and blueshifted emission.
The contours in both top panels start at 3$\sigma$ and increment at steps of
$\pm$3$\sigma$, where $\sigma$\,=\,0.3 mJy beam\pmOne for the top left panel, 
and 
$\sigma$ = 0.4 mJy beam\pmOne (red) and 0.5 mJy beam\pmOne (blue)
for the top right panel.
The crosses denote the
location of the foreground galaxy at $z$\,=\,0.295.
Contours for the first (bottom left) and second (bottom right) moment maps of the \bco line emission
are shown in steps of
50 \kms, and 100 \kms, respectively. 
The synthesis beam size is 4\farcs4 $\times$ 2\farcs0, at PA = 13$\degr$.
\label{fig:CO21mom}}
\end{figure*}

>>>>

\documentclass[]{emulateapj}
\usepackage{amsmath}

% -------------------------------------- Final Edit Markup ----------------------------------
\usepackage[normalem]{ulem}
\usepackage[usenames]{xcolor}
\definecolor{OliveGreen}{cmyk}{0.64,0,0.95,0.40}
\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
\newcommand{\new}[1]{\textcolor{red}{\bf #1}}
\newcommand{\old}[1]{\sout{\textcolor{airforceblue}{\bf #1}}} % {\textcolor{green!55!blue}{\bf #1}} or  {\textcolor{OliveGreen}{\bf #1}}
% -----------------------------------------------------------------------------------------------

\input{preamble}
\slugcomment{To be submitted to the ApJ}
 \makeatletter
% \renewcommand\normalsize{\@setfontsize\normalsize\@xpt{11.56}}
\renewcommand\normalsize{\@setfontsize\normalsize{10.56}{11.4}}
\makeatother

\citestyle{aa}
\shorttitle{Molecular Gas in the Lensed Wet-Merger RXJ1131$-$1231 at $z$\,=\,0.65}
\shortauthors{Leung, Riechers \& Pavesi}

@@

\title{Molecular Gas Dynamics of the strongly-lensed wet-merger RXS J1131$-$1231 at $z$=0.654}
\author{T. K. Daisy Leung, Dominik A. Riechers, and Riccardo Pavesi}
\affil{Department of Astronomy, Space Sciences Building, Cornell University,
Ithaca, NY 14853, USA; \myemail}

@@

\begin{abstract}
We report observations of the \bco and \cco line emission towards
the quadruply-imaged quasar
RXS\,J1131$-$1231 at $z$\,=\,0.654 obtained using the
\pdbi (PdBI) and \carma (CARMA), making this the first spatially resolved
CO study at intermediate redshift.
Our lens modeling shows that the highly asymmetric double-horned line profile
observed in the \bco spectrum is mainly a result of differential lensing, where
the magnification factor varies from $\sim$3 to $\sim$8 across
different kinematic components.
The intrinsically symmetric line profile with a clear velocity
gradient in the source-plane suggests that
the quasar host galaxy is a rotating disk.
In addition, we find a secondary CO-emitting source near RXJ1131 whose location is
consistent with the optically-faint companion reported in previous studies.
The lensing-corrected gas mass of
$M_{\rm gas}$\ssim1.4\E{10}\,\Msun inferred for RXJ1131
and \ssim2\E{9}\,\Msun for its companion correspond
to a gas mass ratio of $\sim$7:1.
The extended CO gas disk in RXJ1131 ($R$$\sim$6\,kpc) yields a dynamical mass
of $M_{\rm dyn}$\ssim8\E{10}\,\Msun, and thus a gas mass fraction of $\sim$19\%,
which is consistent with the trend of decreasing molecular gas fraction since $z$\ssim1.
The lensing-corrected $M^{*}$\ssim3\E{10}\,\Msun,
$M_{\rm dust}$\ssim4\E{8}\,\Msun, $L_{\rm IR}$\ssim1.5\E{12}\,\Lsun, and SFR$_{\rm FIR}$\eq120\pmm63 \sfrU of RXJ1131
together with the source-plane morphology of the gas and the stellar populations
suggest that ongoing \SBs are embedded within the molecular gas reservoir,
and that the nuclear \SBs are likely triggered by
interactions with the companion.
The $M_{\rm BH}$$/$$M_{\rm bulge}$ of $\gtrsim$0.27\% derived for RXJ1131 lies above the
local relation but is consistent with other radio-loud galaxies at this epoch, suggesting
that the co-evolution between RXJ1131 and its host galaxy is different than those of local AGN host galaxies.
\end{abstract}

\keywords{ISM: molecules --
          infrared: galaxies --
          galaxies: mergers --
          galaxies: starburst --
          galaxies: evolution --
          quasars: individual (RXS\,J1131$-$1231)}

%--------------------------------------------------------------------------
%                                Introduction
%--------------------------------------------------------------------------
\section{Introduction}
Studies of galaxy evolution have been focusing on the interplay between
\SF and active galactic nucleus (AGN) activity across cosmic epochs.
It is not well-understood when and how stellar populations of present-day galaxies were assembled,
but it is clear that both the co-moving \SF rate and black hole accretion rate densities
increased substantially since $z$\,$>\,3$ and reached their climax at $z$\ssim$2$, followed by
a rapid decline toward $z$\ssim$0$ \citep[\eg][]{Hopkins06a, Madau14a}.
% Studying the cold gas at the redshift range 0.2 $<$\,$z$\,$<$\,1 of paramount importance to ravel the cause for this decline.
A leading explanation for this decline is related to the evolution of molecular gas content and
\SF efficiency \citep[]{Erb06a, CW13, Walter14a},
but confirmation from direct molecular gas measurements at intermediate redshift
$(0.2\,<\,z\,<\,1)$ is still limited both in number and in the variety of galaxy populations probed ---
only unresolved CO observations have been carried out for $\sim$$70$ ULIRGs up to date \citep{Combes11a, Combes13a}.

% In the current paradigm, it is posited that every nearby, massive galaxy host a supermassive black hole (SMBH), which
% grows by accretion of the cold molecular gas from its host galaxy.
Meanwhile, empirical scaling relations such as the
$M_{\rm BH}$$-$\mbulge relation \citep[e.g.,][]{Magorrian98a, HR04a}
has been established locally, suggesting a co-eval growth between local SMBHs and their host galaxies.
Attempts to extend this relation out to higher redshifts, beyond the peak epoch
of \SF and AGN activity, have been made in recent years.
It has been reported that the $M_{\rm BH}$$/$$M_{\rm bulge}$ ratio of high-$z$ submillimeter galaxies (SMGs) falls below
the local relation \citep[\eg][]{Borys05a,Alexander08a} but that of high-$z$ quasar host galaxies is found
lying above
the relation \citep[\eg][]{McLure06a,Peng06a}.
Stellar mass of quasar host galaxies at $z$\,$>$\,4 constrained by their CO-based dynamical mass
is also too small compared to that based on the local relation \citep{Walter04a, Riechers08a,Coppin08a}.
While it has become clearer that high-$z$ galaxies deviate from the local relation, and among their galaxy populations,
such a picture is still unclear  at intermediate redshift.
This is owed to the difficulties in measuring stellar emission in the host galaxy under the presence of a bright quasar, or alternatively
due to the lack of spatially resolved CO observations to provide dynamical constraints on the stellar mass.
With gravitational lensing, emission of the background quasar and its host galaxy
is magnified and stretched out to different extents, and is being projected onto different locations in the image plane,
allowing spatially resolved measurements their emission in a
relatively uncontaminated fashion.

\defcitealias{Sluse03a}{S03}
The quasar RXS J113151.62$-$123158 (hereafter RXJ1131)
at $z_\textrm{s, QSO}$\eq$0.685$ \citep[hereafter S03]{Sluse03a} is a particularly well-suited source for
studying the evolution of molecular gas properties in quasar host galaxies and the
connection between SMBHs and their host galaxies at intermediate redshift
given its unique lensing configuration.
The stellar emission in the host galaxy of RXJ1131 is lensed into
an Einstein ring of radius $1$\farcs$83$
that is clearly separated from the quadruply imaged quasar emission \citep[hereafter C06]{Claeskens06a}.
We have thus carried out a detailed study of RXJ1131 as presented in this paper.
RXJ1131 is strongly lensed by an elliptical galaxy at $z_\textrm{L}$\eq$0.295$ \citepalias{Sluse03a}.
The moderate mass black hole $M_{\rm BH}$\eq$8$\E{7}\Msun \citep{Sluse12a} %based on H beta line
residing in RXJ1131 is rotating with a high spin parameter \citep[$a$\,$\sim$\,$0.9$;][]{Reis14a}, and
is radiating at an intrinsic bolometric luminosity of $L_{\rm bol, X}$\eq$1.3$\E{45}\, ergs\,s\pmOne \citep{Pooley07a}.

This paper is structured as follows.
In \Sec{obs}, we outline details of the observations and of our data reduction procedures.
In \Sec{HST}, we describe the astrometric corrections
applied to align an {\it HST} image with an VLA radio image.
In \Sec{results}, we report measurements of the \bco and \cco lines and broadband
photometry spanning the optical to radio wavelengths.
In \Sec{anal}, we present dynamical lens modeling of the \bco data and
SED modeling of the host galaxy of RXJ1131 to
derive physical parameters reflecting the nature of this system.
In \Sec{diss}, we discuss the merger stage of RXJ1131,
the velocity offset between the AGN lines and the CO lines,
and the $M_{\rm BH}$$-$\mdyn relation of RXJ1131.
Finally, we summarize the main results and present our conclusions in \Sec{sum}.
We use a concordance $\Lambda$CDM cosmology throughout this paper, with
parameters from the WMAP9 results:
$H_0$ = $69.32$ \kms Mpc\pmOne, $\Omega_{\rm M}$ = $0.29$, and
$\Omega_{\Lambda}$ = $0.71$ \citep{Hinshaw13a}.
% same as LR16a: i.e. use last column (WMAP+eCMB+BAO+H0) of Tab 4 in Hinshaw+13 because that's what astropy use for WMAP9 as cosmo
\defcitealias{Claeskens06a}{C06}

%--------------------------------------------------------------------------
%                          Observations details
%--------------------------------------------------------------------------
\section{Observations} \label{sec:obs}
\subsection{PdBI \bco}
Observations of the \bco rotational line
($\nu_{\rm rest}$\,=\,$230.538$ GHz) redshifted to $\nu_{\rm obs}$\ssim$139.0$\,GHz
were carried out using IRAM \pdbi (PdBI; Program ID: S14BX; PI: D.
Riechers). Two observing runs were carried out on 2014 December 06 and 2015
February 05 under good weather conditions in the C and D array configurations,
respectively. The 2 mm receivers were used to cover the redshifted \bco line
and the underlying continuum emission, employing a correlator setup that provides
an effective bandwidth of 3.6 GHz (dual polarization) and a spectral resolution of 10.0 MHz ($\sim$
$21.5$ \kms). This resulted in 3.75 hours of cumulative six antenna-equivalent on-source
 time after discarding unusable visibility data.
The nearby quasars B1127$-$145 and B1124$-$186 were observed every 22 minutes
for pointing, secondary amplitude, and phase calibration, and B1055$+$018 was
observed as the bandpass calibrator for both tracks.
MWC349 and 3C279 were observed as primary flux calibrators for the C and D
array observations, respectively, yielding calibration accuracy better than 15\%.

The \ncode{gildas} package was used to calibrate and analyze the visibility data.
The calibrated visibility data were imaged and deconvolved using the CLEAN algorithm with ``natural''
weighting. This yields a synthesized clean beam size of 4$\farcs$44 $\times$ 1\farcs95 (PA = 13\degr).
The final rms noise is $\sigma$\,=\,1.45\,mJy\,beam\pmOne
over 10 MHz (21.5\,\kms). The continuum image at $\nu_{\rm cont}\sim$139\,GHz
is produced by averaging over 3.16\,GHz of line-free bandwidth. This
yields an rms noise of 0.082 mJy\,beam$^{-1}$. % see README.md in 04Sep15

@@

The \ncode{miriad} package was used to calibrate the visibility data.
The calibrated visibility data were
imaged and deconvolved using the CLEAN algorithm with ``natural'' weighting. This yields a synthesized clean
beam size of 3\farcs2 $\times$ 1\farcs9 (PA\,=\,8\degr) for the lower sideband
image cube. The final rms noise is $\sigma$ = 13.3 mJy beam$^{-1}$
over a channel width of 25\,MHz. An rms noise of
$\sigma$\,=\,0.83\,mJy\,beam\pmOne is reached by averaging over the
line-free channels in both sidebands.

\subsection{VLA (Archival)} %DONE
Our analysis also uses archival data of the 5\,GHz
radio continuum obtained with the
Very Large Array (VLA; Program ID: AW741; PI: Wucknitz).
Observations were carried out on 2008 December 29 under excellent weather
conditions in the A array configuration for a total of $\sim$7 hours on-source time. The C-band receivers were used with a continuum mode setup,
providing a bandwidth of 50 MHz for the two IF bands with full polarization.
The nearby radio quasar J1130$-$149 was observed every 10 minutes for
pointing, amplitude, and phase calibration, J1331$+$305 was observed as the
primary flux calibrator, and J0319$+$415 was observed as the bandpass
calibrator, yielding $\sim$10\% calibration accuracy.
We used \ncode{aips} to calibrate the visibility data.
The calibrated visibility data were imaged and deconvolved using
the CLEAN algorithm with robust\,=\,0, which
was chosen to obtain a higher resolution image given high SNR.
This yields a synthesized clean
beam size of 0$\farcs$49 $\times$ 0\farcs35 (PA\,=\,0.18$\degr$) and a final
rms noise of $\sigma$ = 13 $\mu$Jy beam\pmOne.

\section{HST astrometry} \label{sec:HST}
We obtained an {\it HST} image taken with
the ACS/Wide Field Camera
using the F555W filter ($V$-band)
from the
Hubble Legacy Archive\footnote{Based on observations
made with the NASA/ESA Hubble Space Telescope, and obtained from the Hubble
Legacy Archive, which is a collaboration between the Space Telescope Science
Institute (STScI/NASA), the Space Telescope European Coordinating Facility
(ST-ECF/ESA) and the Canadian Astronomy Data Centre (CADC/NRC/CSA).}.
The details of the observations can be found
in \citetalias{Claeskens06a}.
We adopt the VLA 5\,GHz map as the
reference coordinate frame to align the optical ($V$-band) image.
We shift the latter to the east by 0\farcs5963 in R.A. and $+$0\farcs8372 in
Dec., which is consistent with the typical astrometric precision (1$^{\prime\prime}-$2$^{\prime\prime}$) of
images from the Hubble Legacy Archive\footnote{http://hla.stsci.edu/hla\_faq.
html}. This astrometric correction is critical to avoid artificial spatial
offsets between different emitting regions and to carry out our lens modeling,
in which the absolute position of the foreground lensing galaxy is based on
its coordinates in the high-resolution optical image.
The VLA image is calibrated using a well-monitored phase
calibrator, with absolute positional accuracy of $\sim$2 mas.
For this reason, the absolute alignment between the VLA image and other
interferometric images reported in this paper are expected to have an astrometric
precision better than 0\farcs1, modulo uncertainties related to the SNR and phase
instability.

%--------------------------------------------------------------------------
%                                Results
%--------------------------------------------------------------------------
\section{Results} \label{sec:results}
\subsection{\bco Emission} \label{sec:CO21}
We detect \bco line emission toward the background source in the PdBI data
at $\gtrsim$\,27$\sigma$ significance, confirming the redshift at $z_{\rm CO}$ =
0.6537\,$\pm$\,0.0005. The emission is spatially and dynamically resolved
with a highly asymmetric double-horned line profile
as shown in \Fig{CO21spec}. Fitting a double Gaussian results in peak
flux densities of 75.3\pmm2.6 mJy and 24.0\pmm2.0 mJy, and a FWHM of
179\pmm9 \kms and 255\pmm28 \kms for the two components, respectively. The peaks are separated by
$\Delta v_{\rm sep}$ = 400\pmm12\,\kms. The total integrated line flux is 24.1\,$\pm$\,2.3 Jy \kms. % see 15May16/

\begin{figure}[htbp]
\centering
\includegraphics[width=0.45\textwidth]{../Figures/SpecCO21_twinx.eps}
\caption{Continuum-subtracted spectrum of \bco emission toward RXJ1131. The velocity scale
is with respect to $z$\,=\,0.6537, which is approximately the line center
considering the asymmetry resulting from differential lensing.
A detailed discussion of this effect is presented in
\Sec{differential}.
 \label{fig:CO21spec}}
\end{figure}

% moment 0 map, highest SNR: CASA chan=125~159 <=> GILDAS 126-160 <=> python 125-160
% - sigma = 0.305 Jy km/s/Beam
%   - note that theoretical sigma is lower = 1.5 * sqrt(160-126+1) * 21.5 ~ 0.2 Jy km/s /beam, due to higher noise in some channels with emission
We construct the zeroth order moment map, a red/blue channel map, and
the first and second moment maps in \Fig{CO21mom}
using the $uv$-continuum subtracted data cube over a velocity range of
$\Delta v$ $\sim$ 750\,\kms.
The higher-order moment maps are produced using
unbinned channel maps with 3$\sigma$ clipping.
The peak flux density is 8.12\pmm0.30 Jy\,\kms\,beam\pmOne
in the intensity map.

The deconvolved source size FWHM from fitting a two-dimensional Gaussian
is 5\farcs1\pmm0\farcs72 $\times$ 3\farcs72\pmm0\farcs66,
and thus, the emission is resolved over $\sim$2.2 beams.
While the lensed emission is not strictly distributed as a two-dimensional
Gaussian;
the fit recovers the line intensity enclosed by the emitting
region, therefore we take this as an estimate on the full extent of the lensed
emission. On the other hand, if we assume that the spatial distribution of
the lensed molecular gas emission is similar to that in the optical to \nir
wavelengths, the lensed emission would be more accurately described by an
annulus, enclosing the partially complete ``Einstein ring'' and
the lensed knots (see \Fig{CO21mom}).

\begin{figure*}[!htbp]
\hspace{0.5em}
\includegraphics[trim=0 10 15 0, clip, width=0.452\textwidth]{../Figures/{F555WCO21_mom0_single.invertedgray}.eps}
\includegraphics[trim=5 -18 0 0, clip, width=0.462\textwidth]{../Figures/F555W_REDBLUE.eps}
\\
\includegraphics[trim=25 16 0 10, clip, width=0.85\textwidth]{../Figures/CO_highOmom_CLIP5sigma}
\vspace{0.1em}
\caption{Top left: overlay of the velocity-integrated \bco emission on the archival {\it HST} $V$-band (F555W)
image.
Top right: same as top left, except the contours are color-coded to represent the red- and blueshifted emission.
The contours in both top panels start at 3$\sigma$ and increment at steps of
$\pm$3$\sigma$, where $\sigma$\,=\,0.3 mJy beam\pmOne for the top left panel,
and
$\sigma$ = 0.4 mJy beam\pmOne (red) and 0.5 mJy beam\pmOne (blue)
for the top right panel.
The crosses denote the
location of the foreground galaxy at $z$\,=\,0.295.
Contours for the first (bottom left) and second (bottom right) moment maps of the \bco line emission
are shown in steps of
50 \kms, and 100 \kms, respectively.
The synthesis beam size is 4\farcs4 $\times$ 2\farcs0, at PA = 13$\degr$.
\label{fig:CO21mom}}
\end{figure*}
